---
title: "Project Report"
author: "Emily Hipp"
date: "10/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Set libraries:
```{r, message=FALSE}
library("tidyverse")
library("gtsummary")
```

Read in data:
```{r, message=FALSE}
df <- read_csv("../DataRaw/lead-iq-01.csv")
# change IQ value of 999 to 99
df["IQ"][df["IQ"] == 999] <- 99
```

IQ levels by location status:
```{r}
df %>%
  ggplot(aes(x = Smelter, y = IQ)) +
  geom_boxplot() +
  ggtitle("Figure 1: IQ levels by location status to lead smelter")
```

IQ summary statistics by location status:
```{r}
df %>%
  tbl_summary(by = Smelter,
              statistic = all_continuous() ~ "{mean} ({sd})"
              ) %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Location**") %>%
  modify_header(label ~ "**Variable**") %>%
  modify_caption("**Table 1. IQ Summary Statistics by Location Status**")
```


```{r include = FALSE}
df.near <- df %>%
  filter(Smelter == "Near")

df.far <- df %>%
  filter(Smelter == "Far")

mean.near <- round(mean(df.near$IQ), 2)
mean.far <- round(mean(df.far$IQ), 2)
```

Figure 1 shows the distribution of IQ values by location status to the lead smelter, while Table 1 shows summary statistics of IQ by location status. The mean IQ of subjects in the "near" group living within 1 mile of the lead smelter is `r mean.near`. The mean IQ of subjects in the "far" group living more than 1 mile from the lead smelter is `r mean.far`. 


